---
title: "Group assignment"
author: "Alima Dzhanybaeva, Jana Hochel, Tasnim Murad"
date: "2022-05-11"
output:
  pdf_document:
         latex_engine: xelatex
  html_document: default
---
```{r include=FALSE}
library(tidyverse)
library(modelsummary)
library(dplyr)
library(knitr)
library(htmlTable)
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```
## 1.1 Overview of the original data.
### The original hotels-europe data includes information on price and features of hotels in 46 European cities and for 2017-2018. The data was downloaded from https://osf.io/yzntm/download.

```{r include=FALSE}
variables <- c('hotel_id','accommodation_type', 'addresscountryname', 'weekend', 'holiday',
               'center1distance', 'starrating', 'guestreviewrating', 'price', 'price_night',
               'year', 'month')
info <- c('Hotel ID', 'Type of accomodation', 'Country', 'Flag, if day is a weekend',
          'Flag, if day is a public holiday', 'Distance from main city center',
          'Number of stars', 'User rating average', 'Price in EUR', 'Number of nights',
          'Year (YYYY)', 'Month (MM)')
type <- c('numeric', 'string', 'string', 'binary', 'binary', 'string',
          'numeric', 'string', 'numeric', 'string', 'numeric', 'numeric')
description <- data.frame(variables, info, type)
```
```{r echo=FALSE}
kable(description, caption = "Description of the main variables in the original dataset")
```

```{r include=FALSE}
df <- read.csv('https://osf.io/yzntm/download')
```
## 1.2 Dataset for 7 Central European countries
### • For our analysis we used 7 Central European countries: Czech Republic, Germany, Italy, Hungary, Austria, Poland, Slovakia.
```{r}
df2 <- df %>% filter(addresscountryname %in% 
                       c('Czech Republic', 'Germany', 'Italy', 
                         'Hungary', 'Austria', 'Poland', 'Slovakia'))

```

### • The folowing manipulations were carried out on the original data to change the existing variables and to add new ones:
* The type of the variables *center1distance* and *guestreviewsrating* were changed from 'string' to 'numeric', and from 'string' to 'factor' for *addresscountryname*. The new variable names are *distance*, *actualrating * and *country* respectively.
```{r include=FALSE}
# guestreviewrating separation
df2 <- separate(df2, guestreviewsrating, ' /', into =
                 c('actualrating', 'rate'))
df2 <- select(df2, -rate)
df2$actualrating <- as.numeric(df2$actualrating)

# Distance
df2 <- separate(df2, center1distance, ' ', into =
                  c('distance', 'miles'))
df2 <- select(df2, -miles)
df2$distance <- as.numeric(df2$distance)

df2$country <- as.factor(df2$addresscountryname)
df2 <- select(df2, -addresscountryname)
```
* The original *accommodationtype* variable was transformed into new *acctype_f* factor variable.
```{r}
df2 <- separate(df2, accommodationtype, '@', into = 
                 c('word', 'acctype'))
df2 <- select(df2, -word)
df2 <- mutate(df2, acctype_f = factor(acctype))
```
* *trueprice* variable was generated by dividing the original *price* variable by *price_night*
```{r}
df2 <- separate(df2, price_night, ' ', into =
                     c('pr','word', 'nights', 'night'))
df2 <- select(df2, -pr)
df2 <- select(df2, -night)
df2 <- select(df2, -word)
df2$nights <- as.numeric(df2$nights)
df2$trueprice <- df2$price/df2$nights
```
* *season* variable was created based on the *month* variable
```{r}
df2 <- df2 %>%
  mutate(season = case_when(month == 12 | month == 1 |month == 2 ~ 'winter',
                            month == 3 | month == 4 |month == 5 ~ 'spring',
                            month == 6 | month == 7 |month == 8 ~ 'summer',
                            month == 9 | month == 10 |month == 11 ~ 'autumn'))
df2$season <- as.factor(df2$season)
```
* With the help of the existing *starrating* variable new *class* variable was generated
```{r}
df2 <- df2 %>%
  mutate(class = case_when(starrating == 1.0 | starrating == 1.5 ~ 'tourist',
                           starrating == 2.0 | starrating == 2.5 ~ 'standard',
                           starrating == 3.0 | starrating == 3.5 ~ 'comfort',
                           starrating == 4.0 | starrating == 4.5 ~ 'first class',
                           starrating == 5.0 ~ 'luxury'))
df2$class <- as.factor(df2$class)
```

### • Quality of the dataset - missing values overview and frequency distribution of numeric values

```{r}
df2 %>% 
  skimr::skim_without_charts() 
```

### • Some value are not blank but include zeros and inf. The next table checks these entries as well as unique values.
```{r echo=FALSE, include=FALSE}
zeros<-funModeling::df_status(df2)
```

```{r echo=FALSE, include=TRUE}
kable(zeros)
```

## • Summary for the main variables in the transformed dataset
## *Numeric variables*
```{r echo=FALSE}
datasummary(distance + starrating + actualrating + trueprice ~ Mean + Min + Max + SD, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```

### There are prominent outliers for *distance* and *true price* variables, therefore, the dataset was transformed, setting the maximum values at the 99% percentile (35 miles and 600 EUR respectively).

```{r echo=FALSE}
df2 <- df2 %>% filter(distance < 35) %>% filter(trueprice < 600)
datasummary(distance + trueprice ~ mean + Min + Max + SD, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```

### • Distance
### 75% of the hotels are located between 0 and 2.4 miles from the main city center. 
### The histogram below shows the distribution of the distances from the city center.

```{r echo=FALSE}
ggplot(data=df2, aes(x=distance)) + geom_histogram(fill='navyblue', color='white', bins = 35) + 
  labs(x='Distance from the city center (in miles)', y='Count')
```

### • Starrating
### Most of hotels from the dataset have 3 and 4 stars.
```{r echo=FALSE}
ggplot(data=df2, aes(x=starrating)) + geom_histogram(fill='navyblue', color='white', bins = 35) + 
  labs(x='Stars', y='Count') + xlim(1,5)
```

### • Actualrating
### As we can see from the histogram below, 80% of the guest reviews are between 3.4 and 4.1.
```{r echo=FALSE, message=FALSE}
ggplot(data=df2, aes(x=actualrating)) + geom_histogram(fill='navyblue', color='white') + 
  labs(x='Guest reviews', y='Count')
```

### • Trueprice
### The distribution of the prices is skewed to the right, therefore, we can conclude that the mean price is greater then the mode.

```{r echo=FALSE, message=FALSE}
ggplot(data=df2, aes(x=trueprice)) + geom_histogram(fill='navyblue', color='white') + 
  labs(x='Price for one night (in EUR)', y='Count') + xlim(0, 600) 
```

### • Year and month
### 22270 records from the dataset are for the year 2017 and 31570 - for 2018.
```{r echo=FALSE}
kable(count(df2, year)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```

### Winter and spring are fully presented in the dataset, whereas there are no records for July, August, September, and October.

```{r echo=FALSE}
kable(count(df2, month)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```

## *Factor variables*

### • Accommodation type
### There are 16 unique accommodation types in the dataset. The most numerous types are "Hotel",  "Apartment", "Guest House",  and "Bed and Breakfast". Together they make up 93% of all records.

```{r echo=FALSE}
kable(count(df2,acctype_f) %>% arrange(desc(n))) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

### • Country
### The countries with the most records are Italy and Germany. Together they make up 66% of all records.
```{r echo=FALSE}
kable(count(df2, country) %>% arrange(desc(n))) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

### • Season and class
### As we observed earlier, the dataset did not include two months for summer and two months for autumn, consequently, spring and winter have the most observations.
```{r echo=FALSE}
kable(count(df2, season) %>% arrange(desc(n))) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

### The most obsevations belong to 'comfort' class, which corresponds to 3 and 3.5 stars, and to 'first class', which corresponds to 4 and 4.5 stars. There are also 15889 missing values. They appeared, because the hotels with 0 stars were not used for the 'class' variable, as the mean price for them is higher than for places with 1, 1.5, 2, and 2.5 stars, which does not make sense. Therefore, these observations were omitted.
```{r echo=FALSE}
kable(count(df2, class)) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

```{r echo=FALSE}
data <- data.frame(
  class=c('comfort', 'first class', 'luxury', 'standard', 'tourist'),
  value=c(17695, 14154, 2307, 4025, 877)
)
data <- data %>% 
  arrange(desc(class)) %>%
  mutate(prop = value / sum(data$value) *100) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )

ggplot(data, aes(x="", y=prop, fill=class)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() + 
  geom_text(aes(y = ypos, label = round(prop, 1)), color = "white", size=5) +
  scale_fill_brewer(palette="Set1") + labs(fill="Class")
```

## *Binary variables*

### • Weekend and holiday
### 35362 records were made on weekends (66%),  and 18478 - on weekdays (34%)

```{r echo=FALSE}
kable(count(df2, weekend)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```

### Only 21% (11546) of all observations were made on holidays.

```{r echo=FALSE}
kable(count(df2, holiday)) %>%
  kableExtra::kable_styling(position = "center", latex_options = "hold_position")
```

## 1.3 Analysis 
### For this project, we wanted to find the best deal among different types of accommodation in Central European countries. 
### Consequently, we formulated the following question: 
* *How does the average price for accommodation change for different values of the following variables: year, country, distance, season, class, weekend?*

## *Findings:*

### • As we can see from the graph below, the price increased for 6 countires, Slovakia being the only case in which the price dropped.
### The biggest jump in price was in Italy. In this country, the price in 2018 increased by 24 euros compared to 2017.

```{r echo=FALSE}
df3 <- aggregate(trueprice ~ country + year, 
                 data=df2, 
                 function(x) { 
                   c(mean_price = mean(x)) 
                 })
ggplot(data = df3, aes(x=year, y=trueprice, fill=country)) + 
  geom_bar(stat='identity') + facet_wrap(~country) +
  geom_text(aes(label = round(trueprice)), vjust = -0.2) +
  scale_x_discrete() + ylim(0, 195) + 
  labs(x='Year(2017,2018)', y='Price for one night (in EUR)', fill='Country')
```

### • The lowest price is 12 EUR per night (Czech Republic),  whereas the lowest average cost is 86 EUR (Poland). On average, Austrian accommodation is the most expensive with a mean price of 158 EUR.

```{r echo=FALSE}
datasummary(trueprice*country ~ Mean + Max + Min, data=df2) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

### • The eminent negative relationship between price and distance from the city center can be tracked only until 7.5 miles. After that point, the dependence of the accommodation cost on the *distance* variable becomes vaguer, as the points are much more dispersed and do not follow a specific pattern. Therefore, we can conclude that on average, the cheapest accommodation is 7.5 miles far from the city center. 

```{r echo=FALSE, message=FALSE}
ggplot(df2, aes(x = distance, y = trueprice))+
  stat_summary_bin(fun = 'mean', bins = 80, 
                   geom = 'point', color = 'red',
                   size = 2) + 
  labs(x='Distance from the city center (in miles)', y='Price for one night (in EUR)') +
  scale_x_continuous(limits = c(0, 35),
                     breaks=seq(from=0, to=35, by=2.5))
```

### • The prices for accommodation in the chosen 7 Central European countries are on average the lowest during autumn and are highest during spring. The difference between the lowest and the highest value is 35 euros.

```{r echo=FALSE}
datasummary(trueprice*season ~ Mean, data=df2) %>%
  kableExtra::kable_styling(latex_options = "HOLD_position")
```

### • As expected, the cheapest class is "tourist" (1 and 1.5 stars) with a mean price of 88 euros. However, the average cost for "standard" (2 and 2.5 stars) falls not that far apart from the best deal with the difference of only 5 euros. Conversely, the "luxury"  (5 stars) class is on average 140 euros more expensive than  "first class" (4 and 4.5 stars). 
### Consequently, the difference between mean prices becomes bigger with the number of stars. 

```{r echo=FALSE}
datasummary(trueprice*class ~ Mean, data = df2) %>%
   kableExtra::kable_styling(latex_options = "hold_position")
```
```{r echo=FALSE, message=FALSE}
ggplot(data = df2, aes(x=reorder(class, +trueprice), y=trueprice, fill = class)) + 
  geom_bar(stat = 'summary') + labs(x='Class', y='Price for 1 night (in EUR)', fill = 'Class')
```

### • The average price for accommodation is 16 euros higher during weekends. Moreover, as we can see from the graph below: (1) the boxplot for weekends is comparatively taller than the one for weekdays, indicating that prices are more dispersed during the end of the week; (2) the upper quartile for weekends is farther from the median than for weekdays, meaning that the 75% percentile is higher.

```{r echo=FALSE}
df2 <- df2 %>%
  mutate(weekend = case_when(weekend == 1 ~'weekend',
                            weekend == 0 ~ 'weekday'))
datasummary(trueprice*weekend ~ Mean, data = df2) %>%
  kableExtra::kable_styling(latex_options = "hold_position")
```
```{r echo=FALSE}
ggplot(data=df2, aes(x=weekend, y=trueprice, group=weekend))+
  geom_violin(trim=FALSE, fill='#A4A4A4', color="darkred")+
  geom_boxplot(width=0.1) + theme_minimal() + 
  labs(x="Weekend", y="Price for one night (in EUR)")
```

## *Summary*
### For the chosen 7 Central European countries (with the maximum value for the *distance* variable set to 35 miles and to 600 euros for *trueprice* variable)  following conclusions on the best deal accross different accommodation types were made:
* Compared to other countries, the average prices in Slovakia do not tend to drastically jump.
* On average, the cheapest deals can be found in Poland.
* The optimal distance from the city center is 7,5 miles.
* Autumn is the best season for finding advantageous offers.
* "Tourist" is the class with the lowest prices, however, "standard" is not much more expensive.
* Accommodation on weekdays is not only cheaper, but the prices are also not as dispersed as on weekends.



